Pricing Games in Heterogeneous Parallel Networks

### 2025-02-09 thesis/slides.tex
Thesis: Pres: Adds pauses and slide flow with final cleanup and fixes

\documentclass{beamer}
\usepackage{tikz}
\usepackage{multicol}
\usepackage[style=authoryear,backend=bibtex]{biblatex}
\addbibresource{ref.bib}

@@

\begin{frame}{\insertsection}
	Selfish routing can model and solve a variety of real-world problems in economics, transportation and more.
	\\[20pt]
	\begin{tikzpicture}[scale=0.9, every node/.style={font=\small,align=center}]
		\draw (0,-0.15) node[below] {Pigou} -- (0,0.15) node[above] {$1920$};
		\draw (0,0) -- (6,0);
		\draw (6,-0.15) node[below] {Beckman\\et al.} -- (6,0.15) node[above] {$1956$};
		\draw (6,0) -- (9,0);
		\draw (9,-0.15) node[below] {Rosenthal} -- (9,0.15) node[above] {$1973$};
		\draw (9,0) -- (10,0);

		\draw (0,-2.65) node[below] {Cole\\et al.} -- (0,-2.35) node[above] {$2003$};
		\draw (0,-2.5) -- (2,-2.5);
		\draw (2,-2.65) node[below] {Acemoglu\\\& Ozdaglar} -- (2,-2.35) node[above] {$2007$};
		\draw (2,-2.5) -- (7.5,-2.5);
		\draw (7.5,-2.65) node[below] {Correa\\et al.} -- (7.5,-2.35) node[above] {$2018$};
		\draw (7.5,-2.5) -- (9,-2.5);
		\draw (9,-2.65) node[below] {Harks\\et al.} -- (9,-2.35) node[above] {$2019$};
		\draw (9,-2.5) -- (10,-2.5);
	\end{tikzpicture}
\end{frame}

@@

\begin{frame}{Preliminaries | Our model}
	We will describe a pricing game as the tuple $\Gm = (N, \ell, \alpha)$ where
	\begin{itemize}
		\item $N = \{1, 2, \cdots, n\}$ the links
		\item $\ell = (\ell_i)_{i \in N}$ the latency functions
		\item $\alpha$ the distribution function
	\end{itemize}
	A homogeneous game can be described as $(N, \ell, 1)$.
\end{frame}

\begin{frame}{Preliminaries | Game equivalence}
	\begin{definition}[Game equivalence]
		Let $\Gm_1, \Gm_2$ be two $n$-link Network Congestion Games.
		$\Gm_1$ and $\Gm_2$ are called \textbf{equivalent} if and only if for all tolls $t \in \R_+^{|N|}$ it holds that $x^{(1)}(t) = x^{(2)}(t)$, with $x^{(1)}(t), x^{(2)}(t)$ being the Nash Equilibria for $t$ in $\Gm_1, \Gm_2$ respectively.
	\end{definition}
	Equivalency means same tolls cause same flow.

@@

@@

	Toll differences are viewed with different scaling.
	\\[15pt]
	\begin{tikzpicture}
		% Title
		\node[above,fixed_height] at (3.7,3.3) {$\alpha(0) = 1$};

		% Axes
		\draw[->] (0,0) -- (7,0) node[right=5,fixed_height] {$c\left(0\right)$};
		\draw[-] (0,0) -- (0,3.2);

@@

		% Y-bars: Tolls
		\filldraw[fill=orange!20] (2/11,2.25) rectangle (35/11,2.75) node[below right] {$3$};
		\filldraw[fill=orange!20] (14/11,1.25) rectangle (36/11,1.75) node[below right] {$2$};
		\filldraw[fill=orange!20] (29/11,0.25) rectangle (40/11,0.75) node[below right] {$1$};
	\end{tikzpicture}
\end{frame}

@@

\begin{frame}{Sensitivity split | Upper bound}
	For the Example we have
	\[\max\{x_1(0), x_2(0)\} = \max\left\{\tfrac23, \tfrac13\right\} = \tfrac23\]
	so if we used a distribution function such as
	\[
		\alpha(p) =
		\begin{cases}
			p + 1 & p \le \tfrac23 \\
			\mathrm{e}^p & p > \tfrac23
		\end{cases}
	\]
	then $\as(t)$ would remain the same for all $t \in \R_+^n$
\end{frame}

@@

\begin{frame}{Pseudo-heterogeneous Pricing Games}
	Consider distribution functions $\alpha(p) = c$ for some $c > 0$.

	Players again all view link costs the same.
	\[c_i(p) = \ell_i(x_i) + c t_i\]
	Behaviour should be homogeneous with tolls "mapped" to $t \rightarrow t/c$.

@@

\begin{frame}{Pseudo-heterogeneous Pricing Games}
	\begin{lemma}
		Let $\Gm_1 = (N, \ell^{(1)}, \alpha)$ where $\forall p \in [0, 1] \; \alpha(p) = c$ for some $c \in \R_+$ and $\Gm_2 = (N, \ell^{(2)}, 1)$ where $\ell^{(2)} = \frac{\ell^{(1)}}{c}$ for the same $c$.
		Then the two games are equivalent.
	\end{lemma}
	\begin{proof}
		For any $t \in \R_+^n$ and any flow $x_i$
		\[
			\sum_{i \in N} \left(\ell_i^{(1)}(x_i^{(2)}(t)) + c t_i\right) \left(x_i^{(2)}(t) - x_i\right) \leq 0
		\]
		$x^{(2)}(t)$ is the unique N.E. for $t$ in $\Gm_1$ so $x^{(1)}(t) = x^{(2)}(t)$.
	\end{proof}
	Costs are proportional.
\end{frame}

\begin{frame}{Pseudo-heterogeneous Pricing Games}
	We will call such games \textit{pseudo-heterogeneous}.

	Uniform increase in money-sensitivity is equivalent to uniform decrease in latencies (lost for general $\alpha$ functions).

	\begin{lemma}
		Let $\Gm_1 = \left(N, \frac{\ell}{c_1}, 1\right), \Gm_2 = \left(N, \frac{\ell}{c_2}, 1\right)$.
		Then $x^{(1)}(c_2 t) = x^{(2)}(c_1 t)$.
	\end{lemma}
	\begin{proof}
		\[
			\begin{aligned}[t]
				&\sum_{i \in N} \left(\frac{1}{c_1}\ell_i(x_i^{(1)}(c_2 t)) + c_2 t_i\right) \left(x_i^{(1)}(c_2 t) - x_i\right) \le 0\\
				\Leftrightarrow &\sum_{i \in N} \left(\frac{1}{c_2}\ell_i(x_i^{(1)}(c_2 t)) + c_1 t_i\right) \left(x_i^{(1)}(c_2 t) - x_i\right) \le 0
			\end{aligned}
		\]
	\end{proof}
\end{frame}

@@

\begin{frame}{Pseudo-heterogeneous Pricing Games}
	\begin{lemma}
		Let $\Gm_1 = (N, \ell, c_1)$ and $\Gm_2 = (N, \ell, c_2)$, $c_1, c_2 > 0$.
		Then
		\begin{enumerate}[$(i)$]
			\item $t^{*(1)} = \frac{c_2}{c_1} t^{*(2)}$
			\item $x^{(1)}(t^{*(1)}) = x^{(2)}(t^{*(2)})$
		\end{enumerate}
	\end{lemma}
	\begin{proof}
		\begin{enumerate}[$(i)$]
			% https://tex.stackexchange.com/a/98399
			\item
			$\begin{aligned}[t]
				t_i^{*(1)} &= B_i^{(1)}(t_{-i}^{*(1)}) \Leftrightarrow \cdots \Leftrightarrow \frac{c_1}{c_2} t_i^{*(1)} = B_i^{(2)}\left(\frac{c_1}{c_2} t_{-i}^{*(1)}\right)
			\end{aligned}$
			\item follows from previous lemma.
			\vspace{-20pt}
		\end{enumerate}
	\end{proof}
	Pricing Nash Equilibrium is always at the same flow.
\end{frame}

@@

\begin{frame}{Zero money-sensitivity}
	Zero money-sensitivity creates issues as it can be exploited.
	\begin{lemma}
		Let $(N, \ell, \alpha)$ where $\alpha(p) = 0$ for $p \in [0, \epsilon]$ for some $\epsilon > 0$.\\
		Then $\exists i \in N$ where $B_i(t_{-i})$ is unbounded for all $t_{-i} \in \R_+^{n - 1}$.
	\end{lemma}
	\begin{proof}
		Links with $\ell_i(0) = \min_{j \in N} \ell_i(0)$ can select $t_i > \max t_{-i}$ and become the lowest-latency link, with players $\alpha(p) = 0$.

@@

\begin{frame}<presentation:0>[noframenumbering]{Step distribution functions}
	Notice that when $x_1$ is fixed then $\as$ isn't and vice-versa.\\
	If the flow is moving, then it does so among players with the same sensitivity type, which is one of the finitely many values $\alpha$ can take.
	If all players of each sensitivity type are using a single link, then $\as$ can take any value between the lowest and highest $\alpha$ value in the upper and lower split respectively.
\end{frame}

@@

\begin{frame}{Step distribution functions}
	If $t^*$ exists, it needs to be at $\as(t^*) = 3$ or $4$.

	Flow will be the same regardless.

	Assume $\Gm^\prime = (N, \ell, 1)$ and we calculate
	\[t^{*\prime} = \left(\frac53, \frac43\right) \qquad x(t^*) = \left(\frac59, \frac49\right)\]

	Since $t_1 > t_2$, link $1$ has the lower split, so the split is at $\alpha\left(\frac59\right) = 4$.

@@

\begin{frame}{Step distribution functions | Pricing N.E. Algorithm}
	\textbf{Summary:} Given $\Gm = ([2], \ell, \alpha), \alpha = (\alpha_k)_{k \in [m]}$
	\begin{enumerate}
		\item Consider $\Gm^\prime = ([2], \ell, 1)$ and calculate $t^{*\prime}, x^\prime(t^{*\prime})$.
		\item Calculate $\as(t^*)$ as the $\alpha$ value of the lower split flow in $x^\prime(t^{*\prime})$
		\item Calculate the candidate $t^* = \tfrac{t^{*\prime}}{\as(t^*)}$ and $\Pi_1(t^*), \Pi_2(t^*)$.
		\item For each $\alpha_k \ne \as(t^*)$, and for each $i \in [2]$
		\begin{enumerate}[$4.1$]
			\item Consider $\Gm_k = ([2], \ell, \alpha_k)$ and calculate $B_{i, k}(t_{-i}^*)$ and tolls $t_{i, k}$ for which $\as(t_i, t_{-i}^*) = \alpha_k$, resulting in two toll ranges for $t_i$.
			\item Search for $\Pi_{i, k}$, the max profit for $i$ within those $t_i$ ranges:
			\begin{enumerate}[$4.2.1$]
				\item If $B_{i, k}(t_{-i}^*)$ is within the $t_i$ ranges, $\Pi_{i, k} = \Pi_i(B_{i, k}(t_{-i}^*), t_{-i}^*)$.
				\item Otherwise, $\Pi_{i, k}$ is max profit among the ranges' edge tolls ($4$).
			\end{enumerate}
			\item If $\Pi_i < \Pi_{i, k}$, then $\Gm$ has NO pricing Nash Equilibrium.
		\end{enumerate}
		\item If all of the above checks fail, return $t^*$.
	\end{enumerate}
	\textbf{Complexity:} $\mathcal{O}(m)$
\end{frame}

@@

\begin{frame}{Future work}
	\textbf{Pseudo-heterogeneous games}
	\begin{itemize}
		\item Connection between similar games with different fixed distribution functions
		\item What happens to flow, tolls, profit when fixed $t_{-i}$?
		\item (Conjectures in the Appendix)
	\end{itemize}

	\textbf{Step distribution functions}
	\begin{itemize}
		\item Conditions for N.E. existence in $2$-link networks
		\item $n$-link networks
		\begin{itemize}
			\item Complexity of calculating existence
		\end{itemize}
	\end{itemize}
\end{frame}

\begin{frame}{Future work}
	\textbf{Continuous distribution functions}
	\begin{itemize}
		\item No gaps!
		\item Function series approximation ($m \rightarrow +\infty$)
		\item Linear $\alpha$ maintains the linear system
		\item Strange behaviour when $0$ is an accumulation point in $\alpha$
	\end{itemize}

	\textbf{Split function for series-parallel links}

>>>>

\documentclass{beamer}
%\documentclass[handout]{beamer}
\usepackage{tikz}
\usepackage{multicol}
\usepackage[style=authoryear,backend=bibtex]{biblatex}
\addbibresource{ref.bib}

@@

\begin{frame}{\insertsection}
	Selfish routing can model and solve a variety of real-world problems in economics, transportation and more.
	\pause
	\\[20pt]
	\begin{tikzpicture}[scale=0.9, every node/.style={font=\small,align=center}]
		\onslide<+->{
			\draw (0,-0.15) node[below] {Pigou} -- (0,0.15) node[above] {$1920$};
		}
		\onslide<+->{
			\draw (0,0) -- (6,0);
			\draw (6,-0.15) node[below] {Beckman\\et al.} -- (6,0.15) node[above] {$1956$};
		}
		\onslide<+->{
			\draw (6,0) -- (9,0);
			\draw (9,-0.15) node[below] {Rosenthal} -- (9,0.15) node[above] {$1973$};
			\draw[->] (9,0) -- (10,0);
		}
		\onslide<+->{
			\draw (0,-2.65) node[below] {Cole\\et al.} -- (0,-2.35) node[above] {$2003$};
		}
		\onslide<+->{
			\draw (0,-2.5) -- (2,-2.5);
			\draw (2,-2.65) node[below] {Acemoglu\\\& Ozdaglar} -- (2,-2.35) node[above] {$2007$};
		}
		\onslide<+->{
			\draw (2,-2.5) -- (7.5,-2.5);
			\draw (7.5,-2.65) node[below] {Correa\\et al.} -- (7.5,-2.35) node[above] {$2018$};
			\draw (7.5,-2.5) -- (9,-2.5);
			\draw (9,-2.65) node[below] {Harks\\et al.} -- (9,-2.35) node[above] {$2019$};
			\draw[->] (9,-2.5) -- (10,-2.5);
		}
	\end{tikzpicture}
\end{frame}

@@

\begin{frame}{Preliminaries | Our model}
	We will describe a pricing game as the tuple $\Gm = (N, \ell, \alpha)$ where
	\begin{itemize}
		\item $N = \{1, 2, \cdots, n\}$ the links
		\item $\ell = (\ell_i)_{i \in N}$ the latency functions
		\item $\alpha$ the distribution function
	\end{itemize}\pause
	A homogeneous game can be described as $(N, \ell, 1)$.
\end{frame}

\begin{frame}{Preliminaries | Game equivalence}
	\begin{definition}[Game equivalence]
		Let $\Gm_1, \Gm_2$ be two $n$-link Network Congestion Games.
		$\Gm_1$ and $\Gm_2$ are called \textbf{equivalent} if and only if for all tolls $t \in \R_+^{|N|}$ it holds that $x^{(1)}(t) = x^{(2)}(t)$, with $x^{(1)}(t), x^{(2)}(t)$ being the Nash Equilibria for $t$ in $\Gm_1, \Gm_2$ respectively.
	\end{definition}
	Equivalency means same tolls cause same flow.\pause

@@

\begin{frame}{Sensitivity split | Heterogeneous networks}
	Latencies are the same for all players.\pause

@@

	Toll differences are viewed with different scaling.\pause[1]
	\\[15pt]
	\begin{tikzpicture}
		% Axes
		\draw[->] (0,0) -- (7,0) node[right=5,fixed_height] {$c$};
		\draw[-] (0,0) -- (0,3.2);

@@

		\onslide<2->{
			% Title
			\node[above,fixed_height] at (3.7,3.3) {$\alpha(0) = 1$};

			% Y-axis label
			\node[right=12,fixed_height] at (7,0) {$(0)$};

			% Y-bars: Tolls
			\filldraw[fill=orange!20] (2/11,2.25) rectangle (35/11,2.75) node[below right] {$3$};
			\filldraw[fill=orange!20] (14/11,1.25) rectangle (36/11,1.75) node[below right] {$2$};
			\filldraw[fill=orange!20] (29/11,0.25) rectangle (40/11,0.75) node[below right] {$1$};
		}
	\end{tikzpicture}
\end{frame}

@@

\begin{frame}{Sensitivity split | Upper bound}
	For the Example we have
	\[\min\{x_1(0), x_2(0)\} = \max\left\{\tfrac23, \tfrac13\right\} = \tfrac13\]\pause
	so if we used a distribution function such as
	\[
		\alpha(p) =
		\begin{cases}
			p + 1 & p \le \tfrac23 \\
			\mathrm{e}^p & p > \tfrac23
		\end{cases}
	\]
	then $\as(t)$ would remain the same for all valid $t \in \R_+^n$.
\end{frame}

@@

\begin{frame}{Pseudo-heterogeneous Pricing Games}
	Consider distribution functions $\alpha(p) = c$ for some $c > 0$.\pause

	Players again all view link costs the same.
	\[c_i(p) = \ell_i(x_i) + c t_i\]
	Behaviour should be homogeneous with tolls "mapped" to $t \rightarrow t/c$.\pause

@@

\begin{frame}{Pseudo-heterogeneous Pricing Games}
	\begin{lemma}
		Let $\Gm_1 = (N, \ell^{(1)}, \alpha)$ where $\forall p \in [0, 1] \; \alpha(p) = c$ for some $c \in \R_+$ and $\Gm_2 = (N, \ell^{(2)}, 1)$ where $\ell^{(2)} = \frac{\ell^{(1)}}{c}$ for the same $c$.
		Then the two games are equivalent.
	\end{lemma}\pause
	\begin{proof}
		For any $t \in \R_+^n$ and any flow $x_i$
		\[
			\sum_{i \in N} \left(\ell_i^{(1)}(x_i^{(2)}(t)) + c t_i\right) \left(x_i^{(2)}(t) - x_i\right) \leq 0
		\]
		$x^{(2)}(t)$ is the unique N.E. for $t$ in $\Gm_1$ so $x^{(1)}(t) = x^{(2)}(t)$.
	\end{proof}
	Costs are proportional.
\end{frame}

\begin{frame}{Pseudo-heterogeneous Pricing Games}
	We will call such games \textit{pseudo-heterogeneous}.\pause

	Uniform increase in money-sensitivity is equivalent to uniform decrease in latencies (lost for general $\alpha$ functions).\pause

	\begin{lemma}
		Let $\Gm_1 = \left(N, \frac{\ell}{c_1}, 1\right), \Gm_2 = \left(N, \frac{\ell}{c_2}, 1\right)$.
		Then $x^{(1)}(c_2 t) = x^{(2)}(c_1 t)$.
	\end{lemma}\pause
	\begin{proof}
		\[
			\begin{aligned}[t]
				&\sum_{i \in N} \left(\frac{1}{c_1}\ell_i(x_i^{(1)}(c_2 t)) + c_2 t_i\right) \left(x_i^{(1)}(c_2 t) - x_i\right) \le 0\\
				\Leftrightarrow &\sum_{i \in N} \left(\frac{1}{c_2}\ell_i(x_i^{(1)}(c_2 t)) + c_1 t_i\right) \left(x_i^{(1)}(c_2 t) - x_i\right) \le 0
			\end{aligned}
		\]
	\end{proof}
\end{frame}

@@

\begin{frame}{Pseudo-heterogeneous Pricing Games}
	\begin{lemma}
		Let $\Gm_1 = (N, \ell, c_1)$ and $\Gm_2 = (N, \ell, c_2)$, $c_1, c_2 > 0$.
		Then
		\begin{enumerate}[$(i)$]
			\item $t^{*(1)} = \frac{c_2}{c_1} t^{*(2)}$
			\item $x^{(1)}(t^{*(1)}) = x^{(2)}(t^{*(2)})$
		\end{enumerate}
	\end{lemma}
	%\begin{proof}
	%	\begin{enumerate}[$(i)$]
	%		% https://tex.stackexchange.com/a/98399
	%		\item
	%		$\begin{aligned}[t]
	%			t_i^{*(1)} &= B_i^{(1)}(t_{-i}^{*(1)}) \Leftrightarrow \cdots \Leftrightarrow \frac{c_1}{c_2} t_i^{*(1)} = B_i^{(2)}\left(\frac{c_1}{c_2} t_{-i}^{*(1)}\right)
	%		\end{aligned}$
	%		\item follows from previous lemma.
	%		\vspace{-20pt}
	%	\end{enumerate}
	%\end{proof}
	Pricing Nash Equilibrium is always at the same flow.
\end{frame}

@@

\begin{frame}{Zero money-sensitivity}
	Zero money-sensitivity creates issues as it can be exploited.
	\begin{lemma}
		Let $(N, \ell, \alpha)$ where $\alpha(p) = 0$ for $p \in [0, \epsilon]$ for some $\epsilon > 0$.\\
		Then $\exists i \in N$ where $B_i(t_{-i})$ is unbounded for all $t_{-i} \in \R_+^{n - 1}$.
	\end{lemma}\pause
	\begin{proof}
		Links with $\ell_i(0) = \min_{j \in N} \ell_i(0)$ can select $t_i > \max t_{-i}$ and become the lowest-latency link, with players $\alpha(p) = 0$.

@@

@@

\begin{frame}{Step distribution functions}
	If $t^*$ exists, it needs to be at $\as(t^*) = 3$ or $4$ (not between).\pause

	Flow will be the same regardless.\pause

	Assume $\Gm^\prime = (N, \ell, 1)$ and we calculate
	\[t^{*\prime} = \left(\frac53, \frac43\right) \qquad x(t^*) = \left(\frac59, \frac49\right)\]\pause

	Since $t_1 > t_2$, link $1$ has the lower split, so the split is at $\alpha\left(\frac59\right) = 4$.\pause

@@

\begin{frame}{Step distribution functions | Pricing N.E. Algorithm}
	\textbf{Summary:} Given $\Gm = ([2], \ell, \alpha), \alpha = (\alpha_k)_{k \in [m]}$
	\begin{enumerate}
		\item Consider $\Gm^\prime = ([2], \ell, 1)$ and calculate $t^{*\prime}, x^\prime(t^{*\prime})$.\pause
		\item Calculate $\as(t^*)$ as the $\alpha$ value of the lower split in $x^\prime(t^{*\prime})$.\pause
		\item Calculate the candidate $t^* = \tfrac{t^{*\prime}}{\as(t^*)}$ and $\Pi_1(t^*), \Pi_2(t^*)$.\pause
		\item For each $\alpha_k \ne \as(t^*)$, and for each $i \in [2]$:\pause
		\begin{enumerate}[$4.1$]
			\item Consider $\Gm_k = ([2], \ell, \alpha_k)$ and calculate $B_{i, k}(t_{-i}^*)$ and tolls $t_{i, k}$ for which $\as(t_i, t_{-i}^*) = \alpha_k$, resulting in two toll ranges for $t_i$.\pause
			\item Search for $\Pi_{i, k}$, the max profit for $i$ within those $t_i$ ranges:\pause
			\begin{enumerate}[$4.2.1$]
				\item If $B_{i, k}(t_{-i}^*)$ is within the $t_i$ ranges, $\Pi_{i, k} = \Pi_i(B_{i, k}(t_{-i}^*), t_{-i}^*)$.\pause
				\item Otherwise, $\Pi_{i, k}$ is max profit among the ranges' edge tolls ($4$).
			\end{enumerate}\pause
			\item If $\Pi_i < \Pi_{i, k}$, then $\Gm$ has NO pricing Nash Equilibrium.
		\end{enumerate}\pause
		\item If all of the above checks fail, return $t^*$.
	\end{enumerate}\pause
	\textbf{Complexity:} $\mathcal{O}(m)$
\end{frame}

@@

\begin{frame}{Future work}
	\textbf{Pseudo-heterogeneous games}\pause
	\begin{itemize}
		\item Connection between similar games with different fixed distribution functions\pause
		\item What happens to flow, tolls, profit when fixed $t_{-i}$?\pause
		\item (Conjectures in the Appendix)
	\end{itemize}\pause

	\textbf{Step distribution functions}\pause
	\begin{itemize}
		\item Conditions for N.E. existence in $2$-link networks\pause
		\item $n$-link networks\pause
		\begin{itemize}
			\item Complexity of calculating existence
		\end{itemize}
	\end{itemize}
\end{frame}

\begin{frame}{Future work}
	\textbf{Continuous distribution functions}\pause
	\begin{itemize}
		\item No gaps!\pause
		\item Function series approximation ($m \rightarrow +\infty$)\pause
		\item Linear $\alpha$ maintains the linear system\pause
		\item Strange behaviour when $0$ is an accumulation point in $\alpha$
	\end{itemize}\pause

	\textbf{Split function for series-parallel links}\pause
